Deep Learning

### 2015-03-25 deepreview.tex
implemented most of Bernhard's suggestions (ref 1)

% classical extensions of simple classifiers
There have been numerous attempts, going back to the early 1960s, to
increase the flexibility of classifiers so as to allow them to
separate categories from one another by more complex surfaces than
flat hyperplanes. A popular example is the class of kernel methods,
particularly the Support Vector Machine~\citep{Boser92} which
represents inputs by a vector whose components are similarities to
each of the training samples computed through a so-called {\em kernel
  function}. While these methods are good general-purpose classifiers,
they do not perform well for tasks such as image or speech recognition
in which the input space is very high-dimensional and the input-output
mapping is so complex that computing similarities using a simple
kernel function is ineffective. Designing feature extractors that are
sensitive to the relevant information and invariant to the irrelevant
one is what generations of computer vision and pattern recognition
researchers have spent their careers on.

@@

The idea that this could be done, and that it could work at all, took
a surprisingly long time to emerge. It was discovered independently by
everal different groups in the 1970s and
1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
in the late 1980s, spurring the second wave of interest for neural
networks. But for reasons that historians of science will have to
explain, the method was the subject of scorn by the ML community
between the late 1990s and the late 2000s.

@@

If ConvNets existed in the 1990s, why did they become so popular and
dominant only now? Certainly, the prejudice of the ML and vision
communities against neural nets played a role. But a number of recent
ideas and technologies enabled the success. First, ConvNets strive
when training data is abundant, and large labeled image datasets such
as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
the appearance of GPU cards capable of over 1Tflops, together with
GPU-optimized software for ConvNet made it possible to train these
large networks in a few weeks (now days) of computer time. Third, the
use of rectifying non-linearities instead of sigmoids allowed to train
much deeper networks than previously possible without unsupervised
pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
the ``drop out'' regularization method~\citep{Srivastava14} was shown
to mitigate the effects of over-fitting in very large networks.

@@

The hidden layers of a multi-layer neural network learn to represent
the network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}.  Each context word is
presented to the network as a one-of-N vector, i.e., in which one
component has a value of $1$ and the rest are $0$. The output layer
has one unit per word and is forced to produce a probability
distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{e^{z_j}}{\sum_i e^{z_i}}
\end{equation}
where $z_j$ is the total input received by output unit $j$. 
\iffalse
The hidden
layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
\begin{equation}
y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
 y_j = \max(0, z_j).
\end{equation}
\fi

The presence of a particular word at the $i$-th position in the input
of the network sets the value of specialized hidden units to values
that correspond to the representation of that word. Different words
create a different pattern of activations, or word-vector, illustrated
in Figure~\ref{fig:word-embeddings}.  In a language model, the other
layers of the network learn to convert the input word-vectors into an
output word-vector for the predicted next-word, which can be used to
predict the probability for any word in the vocabulary to appear as
the next word.  The network learns word-vectors that contain many
active components each of which can be interpreted as a separate
feature of the word. This was first demonstrated using sequences of
three words, such as (Colin, Mother, Victoria), derived from two
family trees~\citep{RHW}. The learned word vector for Colin contained
``semantic'' features representing which family tree Colin was in,
what branch of that tree he was in, and what generation he was from.
%% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
%% END IN ORDER NOT TO SOUND STILTED.
These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

The family tree demonstration used a small, clean dataset in which the
regularities could be captured by micro-rules that were never violated
and the features were easy to interpret, but learned word-vectors also
work very well when the word sequences come from a large corpus of
real text and the individual micro-rules are
unreliable~\citep{BenDucVin01-short}. When trained to predict the next
word in a news story, for example, the learned word vectors for
Tuesday and Wednesday are very similar, as are the word vectors for
Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
put in the same category, but rather that they share many semantic and
grammatical attributes, so that what the neural network learns from
one day of the week can often be generalized to the other, while
individual days remain different (for example one imagines that the
neural network also learns some features that distinguishes week-end
days and weekdays, while Saturday and Sunday might have special roles
in some religions, etc.).  Such representations are called {\em
  distributed representations} because its elements (the features) are
not mutually exclusive and there exists a rich and vast set of
possible configurations of values of these features, corresponding to
the variations seen in the observed data.  These word-vectors are thus
{\em learned features} that were not determined ahead of time by
humans, but automatically discovered by the neural network.  Vector
representations of words learned from text are now very widely used in
natural language
applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
  Mikolov-et-al-NIPS2013,Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When backpropagation was first introduced, its most exciting use was
for training recurrent neural networks. 
For tasks that involve sequential inputs, such as speech and language,
it is often better to use recurrent neural networks, illustrated in
Figure~\ref{fig:rnn}.  Recurrent neural networks process an input
sequence one element at a time, maintaining in their hidden units a
``state vector'' that implicitly contains information about the
history all the past elements of the sequence.  
When we consider the outputs of
the hidden units at different discrete time steps as if they were the
outputs of different neurons in a deep multi-layer network (right of
Figure~\ref{fig:rnn}), it becomes
clear how we can apply backpropagation to train recurrent
networks. The the gradient with respect to a recurrent weight,
$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at
time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
input sequence is just the sum of its gradients at each time-step.

@@

Thanks to advances in their architecture and ways of training them, recurrent networks 
have been found to be very good at predicting the next character in
text~\citep{Sutskever-et-al-ICML2011} or the next word in a
sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a
time, for example, an English ``encoder'' network can be trained so
that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state of (or as extra
input to) a jointly trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen from this
distribution and provided as input to the decoder network it will then
output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

\begin{figure}[htp*]
\fbox{
\begin{minipage}{\textwidth}
%\centerline{
%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
%}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{
%Top left: graph of feedforward computations in a multi-layer
%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
%one output layer (out), a target $y$, and an error that compares the output
%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
%the parameters of the neural network.\newline
%Top right: back-propagation through the same multi-layer neural network
%(downward arrows compute the partial derivative of the error with respect
%to every node, i.e., the units value and the parameters).\newline
A recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the sampled outputs (e.g. words) can be generated according to an output 
distribution (the probability of the next word, for every possible next word) 
and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:rnn}
\end{minipage}
}
\end{figure}

@@

Memory Networks can be trained to keep track of the state of the world
when they are told a story, so that they can answer questions about
the state of the world at the end of the story. Following a sequence
of actions in a virtual world described in words, in a scenario
similar to a text adventure game, the network is shown an
automatically0generate question and taught to produce the
corresponding answer (also generated automatically).  In one test
example, the network is shown a 15-sentence version of the Lord of The
Rings and correctly answers questions like ``where is Frodo now?''.
~\citep{weston-memorynet-2014}

@@

In a somewhat related work, a network is trained to evaluate whether a
particular mathematical transformation of a formula is lilely to lead
to a correct mathematical identity~\cite{zaremba-nips-2014}. The
system was able to produce efficient identities for matrix expressions
whose formula fill several pages.

@@

Learning which action will maximize the total future reward given a state
of the world is difficult when the rewards are delayed and the world has a
very large number of possible states. One impressive recent development is
a single system that can learn to play many different video
games~\citep{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
map from the pixels on the screen at several time steps to the
expected future reward of each possible action and its only inputs are the
pixels and the current score in the game. The learning works by minimizing
the difference between the system's current estimate of future reward and a
noisy but less biased estimate that is obtained by using the same deep
network to estimate the future reward of the next state, where the next
state is chosen stochastically based on the predicted future rewards of all
the currently legal actions.

@@

It also took an inordinate amount of time for the community to adopt
stochastic gradient methods for large-scale machine learning. key
theoretical and practical work showed that, while SGD has terrible
asymptotic convergence properties, it is extremely efficient for the
kind of approximate optimization required in ML
applications~\citep{bottou-bousquet-2008-small}. But when trainnig a
large ConvNet can take days or weeks on the fastest GPUs, one is eager
to discover more efficient optimization methods that use the curvature
information the past update directions~\citep{lecun-98b,Montavon2012}.

The parallelization of large neural nets over a single GPU has
required to use ``mini-batches'' of samples over which the gradients
are averaged. The purpose of this is to saturate the GPU. But one big
unsolved question is how to scale up the training of very large neural
nets over multiple GPU card across multiple machines. As of today, this
question is unresolved.

Interest in deep feedforward networks was revived around
2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
by the introduction of unsupervised learning procedures that could
create layers of feature detectors without requiring labeled data. The
objective in learning each layer of feature detectors was to be able
to reconstruct or model the activities of feature detectors (or raw
inputs) in the layer below.  By ``pre-training'' several layers of
progressively more complex feature detectors using this reconstruction
objective, the weights of a deep network could be initialized to
sensible values.  A final layer of output units could then be added to
the top of the network and the whole deep system could be fine-tuned
using standard backpropagation
\citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}.
This worked remarkably well for recognizing handwritten digits or for
detecting pedestrians, especially when the amount of labeled data was
very limited~\citep{Erhan+al-2010-small,sermanet-cvpr-13}.

>>>>

% classical extensions of simple classifiers
% YB: I propose to mostly remove the paragraph below (but keep the Scholkopf book ref)
There have been numerous attempts, going back to the early 1960s, to
increase the flexibility of classifiers so as to allow them to
separate categories from one another by more complex surfaces than
flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
particularly the Support Vector Machine~\citep{Boser92} which
represents inputs by a vector whose components are similarities to
each of the training samples computed through a so-called {\em kernel
  function}. While these methods are good general-purpose classifiers,
they do not perform well for tasks such as image or speech recognition
in which the input space is very high-dimensional and the input-output
mapping is so complex that computing similarities using a simple
kernel function is ineffective. Designing feature extractors that are
sensitive to the relevant information and invariant to the irrelevant
one is what generations of computer vision and pattern recognition
researchers have spent their careers on.

@@

The idea that this could be done, and that it could work at all, took
a surprisingly long time to emerge. It was discovered independently by
several different groups in the 1970s and
1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
in the late 1980s, spurring the second wave of interest for neural
networks. But for reasons that historians of science will have to
explain, the method was the subject of scorn by the ML community
between the late 1990s and the late 2000s.

@@

If ConvNets existed in the 1990s, why did they become so popular and
dominant only now? The idea of convexity of the objective function grew
as an attractive property of learning algorithms, thanks to the success
of kernel machines, which made neural nets (that involve a non-convex
training criterion) less appealing. As discussed below, it is only
in 2014 that theoretical and empirical evidence have suggested that local
minima are probably not as much of a problem as was initially thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
Much before that realization, though, a number of recent
ideas and technologies enabled the success. First, supervised ConvNets work best
when training data is abundant, and large labeled image datasets such
as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
the appearance of GPU cards capable of over 1Tflops, together with
GPU-optimized software for ConvNet made it possible to train these
large networks in a few weeks (now days) of computer time. Third, the
use of rectifying non-linearities instead of sigmoids allowed to train
much deeper networks than previously possible without unsupervised
pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
the ``drop out'' regularization method~\citep{Srivastava14} was shown
to mitigate the effects of over-fitting in very large networks.

@@

The hidden layers of a multi-layer neural network learn to represent
the network's inputs in a way that makes it easy to predict the target
outputs. This is nicely demonstrated by training a multi-layer neural
network to predict the next word in a sequence from a local context of
earlier words~\citep{BenDucVin01-short}.  Each context word is
presented to the network as a one-of-N vector, i.e., in which one
component has a value of $1$ and the rest are $0$. 
%The output layer
%has one unit per word and is forced to produce a probability
%distribution over words by using the ``softmax'' function:
%\begin{equation}
%p_j = \frac{e^{z_j}}{\sum_i e^{z_i}}
%\end{equation}
%where $z_j$ is the total input received by output unit $j$. 
\iffalse
The hidden
layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
\begin{equation}
y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
 y_j = \max(0, z_j).
\end{equation}
%
%The presence of a particular word at the $i$-th position in the input
%of the network sets the value of specialized hidden units to values
%that correspond to the representation of that word. 
In the first layer, each word
creates a different pattern of activations, or word-vector, illustrated
in Figure~\ref{fig:word-embeddings}.  In a language model, the other
layers of the network learn to convert the input word-vectors into an
output word-vector for the predicted next-word, which can be used to
predict the probability for any word in the vocabulary to appear as
the next word.  The network learns word-vectors that contain many
active components each of which can be interpreted as a separate
feature of the word, as was first demonstrated by~\citet{RHW} in the context
of learning distributed representations for symbols.
%. This was first demonstrated using sequences of
%three words, such as (Colin, Mother, Victoria), derived from two
%family trees~\citep{RHW}. The learned word vector for Colin contained
%``semantic'' features representing which family tree Colin was in,
%what branch of that tree he was in, and what generation he was from.
%% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
%% END IN ORDER NOT TO SOUND STILTED.
These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  
%For example, given Colin's
%generation and the learned feature of Mother that the output should be one
%generation higher than the input, we can already restrict the answer to
%people in a specific generation. The ability of deep learning to discover
%many such micro-rules allows it to generalize well to novel combinations of
%individually familiar context words.
%
%The family tree demonstration used a small, clean dataset in which the
%regularities could be captured by micro-rules that were never violated
%and the features were easy to interpret, 
Learning word-vectors turned out to also
work very well when the word sequences come from a large corpus of
real text and the individual micro-rules are
unreliable~\citep{BenDucVin01-short}. When trained to predict the next
word in a news story, for example, the learned word vectors for
Tuesday and Wednesday are very similar, as are the word vectors for
Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
put in the same category, but rather that they share many semantic and
grammatical attributes, so that what the neural network learns from
one day of the week can often be generalized to the other, while
individual days remain different (for example one imagines that the
neural network also learns some features that distinguishes week-end
days and weekdays, while Saturday and Sunday might have special roles
in some religions, etc.).  Such representations are called {\em
  distributed representations} because its elements (the features) are
not mutually exclusive and there exists a rich and vast set of
possible configurations of values of these features, corresponding to
the variations seen in the observed data.  These word-vectors are thus
{\em learned features} that were not determined ahead of time by
humans, but automatically discovered by the neural network.  Vector
representations of words learned from text are now very widely used in
natural language
applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
  Mikolov-et-al-NIPS2013,Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When backpropagation was first introduced, its most exciting use was
for training recurrent neural networks. 
For tasks that involve sequential inputs, such as speech and language,
it is often better to use recurrent neural networks, illustrated in
Figure~\ref{fig:rnn}.  Recurrent neural networks process an input
sequence one element at a time, maintaining in their hidden units a
``state vector'' that implicitly contains information about the
history of all the past elements of the sequence.  
When we consider the outputs of
the hidden units at different discrete time steps as if they were the
outputs of different neurons in a deep multi-layer network (right of
Figure~\ref{fig:rnn}), it becomes
clear how we can apply backpropagation to train recurrent
networks. The gradient with respect to a recurrent weight,
$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at
time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
input sequence is just the sum of its gradients at each time-step.

@@

Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
and ways of training them~\citep{Sutskever-thesis2012}, recurrent networks 
have been found to be very good at predicting the next character in
text~\citep{Sutskever-et-al-ICML2011} or the next word in a
sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a
time, for example, an English ``encoder'' network can be trained so
that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state of (or as extra
input to) a jointly trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen from this
distribution and provided as input to the decoder network it will then
output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

\begin{figure}[htp*]
\fbox{
\begin{minipage}{\textwidth}
%\centerline{
%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
%}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{
%Top left: graph of feedforward computations in a multi-layer
%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
%one output layer (out), a target $y$, and an error that compares the output
%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
%the parameters of the neural network.\newline
%Top right: back-propagation through the same multi-layer neural network
%(downward arrows compute the partial derivative of the error with respect
%to every node, i.e., the units value and the parameters).\newline
A recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the network can generate a sequence of outputs (e.g. words), each of which
is used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:rnn}
\end{minipage}
}
\end{figure}

@@

Memory Networks can be trained to keep track of the state of the world
when they are told a story, so that they can answer questions about
the state of the world at the end of the story. Following a sequence
of actions in a virtual world described in words, in a scenario
similar to a text adventure game, the network is shown an
automatically generate question and taught to produce the
corresponding answer (also generated automatically).  In one test
example, the network is shown a 15-sentence version of the Lord of The
Rings and correctly answers questions like ``where is Frodo now?''.
~\citep{weston-memorynet-2014}

@@

% Bernhard was not impressed, so maybe just cut this off
%In a somewhat related work, a network is trained to evaluate whether a
%particular mathematical transformation of a formula is lilely to lead
%to a correct mathematical identity~\citep{zaremba-nips-2014}. The
%system was able to produce efficient identities for matrix expressions
%whose formula fill several pages.

@@

Learning which action will maximize the total future reward given a state
of the world is difficult when the rewards are delayed and the world has a
very large number of possible states. One impressive recent development is
a single system that can learn to play many different video
games~\citep{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
map from the pixels on the screen at several time steps to the
expected future reward of each possible action (assuming some policy is followed
in the future) and its only inputs are the
pixels and the current score in the game. The learning works by minimizing
the difference between the system's current estimate of future reward and a
noisy but less biased estimate that is obtained by using the same deep
network to estimate the future reward of the next state, where the next
state is chosen stochastically based on the predicted future rewards of all
the currently legal actions.

@@

It also took an inordinate amount of time for the community to adopt
stochastic gradient methods for large-scale machine learning. Key
theoretical and practical work showed that, while SGD has terrible
asymptotic convergence properties, it is extremely efficient for the
kind of approximate optimization required in ML
applications~\citep{bottou-bousquet-2008-small}. But when training a
large ConvNet can take days or weeks on the fastest GPUs, one is eager
to discover more efficient optimization methods that use the curvature
information the past update directions~\citep{lecun-98b,Montavon2012}.

The parallelization of large neural nets over a single GPU has
required to use ``mini-batches'' of samples over which the gradients
are averaged. The purpose of this is to saturate the GPU. But one big
unsolved question is how to scale up the training of very large neural
nets over multiple GPU cards across multiple machines. As of today, this
question is unresolved.

Interest in deep feedforward networks was revived around
2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
thanks to the effort of a forward-looking Canadian funding agency 
(the Canadian Institute for Advanced Research, or CIFAR), with
the introduction of unsupervised learning procedures that could
create layers of feature detectors without requiring labeled data. The
objective in learning each layer of feature detectors was to be able
to reconstruct or model the activities of feature detectors (or raw
inputs) in the layer below.  By ``pre-training'' several layers of
progressively more complex feature detectors using this reconstruction
objective, the weights of a deep network could be initialized to
sensible values.  A final layer of output units could then be added to
the top of the network and the whole deep system could be fine-tuned
using standard backpropagation
\citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}.
This worked remarkably well for recognizing handwritten digits or for
detecting pedestrians, especially when the amount of labeled data was
very limited~\citep{Erhan+al-2010-small,sermanet-cvpr-13}.
